% recovery-paper.tex
% Recovering Write-Protected NVMe SSDs via USB Bridge XRAM Command Injection
% J. Sullivan, 2026

\documentclass[10pt,twocolumn]{article}

% ── Packages ──────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[margin=0.75in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{fancyhdr}

% ── Colors ────────────────────────────────────────────────────────────
\definecolor{codebg}{HTML}{F5F5F5}
\definecolor{codeframe}{HTML}{CCCCCC}
\definecolor{codecomment}{HTML}{6A737D}
\definecolor{codestring}{HTML}{032F62}
\definecolor{codekeyword}{HTML}{D73A49}
\definecolor{hexblue}{HTML}{0366D6}
\definecolor{successgreen}{HTML}{28A745}
\definecolor{failred}{HTML}{CB2431}

% ── Listings Configuration ────────────────────────────────────────────
\lstset{
  basicstyle=\ttfamily\footnotesize,
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codeframe},
  breaklines=true,
  breakatwhitespace=false,
  columns=fullflexible,
  keepspaces=true,
  tabsize=2,
  showstringspaces=false,
  commentstyle=\color{codecomment}\itshape,
  keywordstyle=\color{codekeyword}\bfseries,
  stringstyle=\color{codestring},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=5pt,
  xleftmargin=1em,
  framexleftmargin=1em,
  captionpos=b,
}

\lstdefinestyle{hex}{
  basicstyle=\ttfamily\footnotesize\color{hexblue},
  numbers=none,
  xleftmargin=0.5em,
  framexleftmargin=0.5em,
}

\lstdefinestyle{shell}{
  language=bash,
  morekeywords={sudo,zig,modprobe,rmmod,echo,dd,xxd,wipefs},
}

\lstdefinestyle{zig}{
  morekeywords={pub,const,fn,struct,return,var,if,else,while,for,try,catch,switch,break,defer,errdefer,and,or,not,true,false,null,undefined,error,void,u8,u16,u32,u64,i32,i64,usize,bool,comptime},
  morecomment=[l]{//},
  morestring=[b]{"},
}

% ── Hyperref Configuration ────────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=hexblue,
  urlcolor=hexblue,
  pdftitle={Recovering Write-Protected NVMe SSDs via USB Bridge XRAM Command Injection},
  pdfauthor={J. Sullivan},
}

% ── Headers ───────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\footnotesize Sullivan -- NVMe Recovery via XRAM Injection}
\fancyhead[R]{\footnotesize\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ── Title ─────────────────────────────────────────────────────────────
\title{%
  \textbf{Recovering Write-Protected NVMe SSDs via\\
  USB Bridge XRAM Command Injection}
}
\author{
  J.~Sullivan\\
  \texttt{github.com/jsullivan2/hiberpower-ntfs}
}
\date{March 2026}

% ══════════════════════════════════════════════════════════════════════
\begin{document}
\maketitle

% ── Abstract ──────────────────────────────────────────────────────────
\begin{abstract}
We present a novel technique for recovering NVMe SSDs that have entered
firmware-level write-protection mode following Flash Translation Layer (FTL)
corruption. When an NVMe SSD is connected through a USB bridge, the bridge
firmware's opcode whitelist typically blocks all recovery commands (Format NVM,
Sanitize, Security Send/Receive), making the drive appear unrecoverable via
USB. We demonstrate that the ASMedia ASM2362 USB-to-NVMe bridge chip exposes
vendor-specific SCSI commands (0xE4/0xE5) providing direct read/write access
to its internal XRAM, where the NVMe Admin Submission Queue resides. By
writing NVMe command structures directly into this queue memory and ringing
the PCIe doorbell via TLP generation, we bypass the bridge's opcode whitelist
entirely. Using this method, we successfully recovered a Phison PS5012-E12
controller (Silicon Power 256GB) that had been in read-only mode for over a
month, executing a Sanitize Block Erase command that standard tools could not
deliver. The technique generalizes to other ASMedia bridge variants and
potentially to other bridge chipsets with exposed internal memory. We provide
an open-source implementation comprising approximately 4,000 lines of Zig
with 35 unit tests.
\end{abstract}

\tableofcontents
\vspace{1em}

% ══════════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}

NVMe solid-state drives can enter a firmware-controlled read-only mode when
internal errors corrupt the Flash Translation Layer (FTL). This protective
mechanism preserves existing data when the controller detects that further
write operations could lead to data loss or controller instability. While the
NVMe specification~\cite{nvme-spec-2.0} provides recovery commands---Format
NVM (opcode \texttt{0x80}) and Sanitize (opcode \texttt{0x84})---these
commands require direct admin queue access that is not always available.

The FTL is the critical firmware component that maps logical block addresses
(LBAs) to physical NAND flash locations. It manages wear leveling, garbage
collection, and bad block management. When power is lost unexpectedly during
FTL metadata updates---such as during a Windows hibernate write followed by
power failure---the mapping tables can become inconsistent. The controller
then enters a protective read-only mode rather than risk corrupting the
underlying NAND data with stale or incorrect address mappings.

\subsection{The USB Bridge Problem}

A common scenario occurs when the affected SSD is housed in a USB enclosure.
USB-to-NVMe bridge chips translate between the SCSI command set (used over
USB) and NVMe admin commands. However, these bridges implement firmware-level
whitelists that restrict which NVMe opcodes can pass through the translation
layer. The ASMedia ASM2362---one of the most widely deployed USB 3.1 Gen 2
NVMe bridges---only forwards \texttt{Identify} (opcode \texttt{0x06}) and
\texttt{Get Log Page} (opcode \texttt{0x02}) through its proprietary 0xE6
passthrough CDB. All other NVMe admin commands, including those necessary for
drive recovery, are silently dropped by the bridge firmware. The SCSI layer
returns success status, creating the illusion that the command was delivered
when in fact the NVMe controller never received it.

This whitelist exists for legitimate reasons: destructive commands like
Format and Sanitize could cause data loss if issued accidentally, and the
bridge firmware designers chose to restrict passthrough to read-only
diagnostics. However, this restriction also prevents legitimate recovery
operations, leaving users with three unpleasant options: (1)~remove the SSD
from its enclosure and connect it to an M.2 slot or PCIe adapter,
(2)~use Windows-only vendor tools that may or may not support the specific
failure mode, or (3)~pay \$300--\$1,500 for professional data recovery
services using specialized hardware like the PC-3000 SSD.

\subsection{The Bypass: XRAM Injection}

This paper describes a fourth option that we developed and successfully used
to recover a Silicon Power 256GB NVMe SSD (Phison PS5012-E12 controller) that
had been in write-protected mode since January 2026 following a Windows
hibernate corruption event combined with an unexpected power loss. The core
insight is that the ASM2362 exposes vendor-specific SCSI commands
(\texttt{0xE4} for XRAM read, \texttt{0xE5} for XRAM write) that provide
direct access to the bridge chip's internal memory, where the NVMe Admin
Submission Queue (Admin SQ) resides. By writing a 64-byte NVMe Submission
Queue Entry (SQE) directly into the Admin SQ via XRAM and then ringing the
SQ Tail Doorbell via PCIe Transaction Layer Packet (TLP) generation, we can
inject arbitrary NVMe admin commands that bypass the bridge firmware's opcode
filter entirely.

The technique is conceptually analogous to a buffer overflow exploit: the
bridge firmware enforces its whitelist in the ``normal'' command path
(0xE6), but it fails to protect the underlying memory structures (XRAM)
where those commands ultimately reside. By writing directly to the queue
memory, we sidestep the enforcement layer.

\subsection{Contributions}

This work makes the following contributions:

\begin{enumerate}[leftmargin=*]
  \item \textbf{XRAM injection technique}: A method for injecting arbitrary
    NVMe admin commands through a USB bridge by writing directly to the
    bridge's internal XRAM, bypassing the firmware opcode whitelist.
  \item \textbf{PCIe TLP doorbell mechanism}: Documentation of how to ring
    the NVMe Admin SQ Tail Doorbell via the ASM2362's PCIe TLP generation
    engine without disconnecting USB.
  \item \textbf{Practical recovery procedure}: A step-by-step procedure that
    successfully recovered a write-protected Phison PS5012-E12 SSD through
    USB, including the discovery that the Admin SQ depth is 4 (not 8) and
    that only Sanitize Block Erase (not Format NVM or Crypto Erase) clears
    the FTL write protection on this controller.
  \item \textbf{Systematic failure catalog}: A detailed catalog of failed
    approaches, documenting each dead end to save future researchers
    from repeating them.
  \item \textbf{Open-source implementation}: A Zig-based tool
    (\texttt{asm2362-tool}) comprising approximately 4,000 lines of code
    with 35 unit tests implementing the complete injection pipeline.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{Background}
\label{sec:background}

\subsection{NVMe Architecture}
\label{sec:nvme-arch}

The NVMe (Non-Volatile Memory Express) specification~\cite{nvme-spec-2.0}
defines a register-level interface between the host and NVMe controller
optimized for non-volatile memory over PCIe. At its core, NVMe uses a
paired-queue model: Submission Queues (SQ) for sending commands from host to
controller, and Completion Queues (CQ) for receiving results. Queue pair 0
is the Admin Queue pair, used for device management; queue pairs 1--N are
I/O Queue pairs for data transfer.

\subsubsection{Submission Queue Entry Format}

Each Submission Queue Entry (SQE) is 64 bytes, organized as 16 DWORDs
(32-bit words, little-endian). Table~\ref{tab:sqe-format} shows the complete
SQE structure. The first DWORD (CDW0) contains the opcode in bits [7:0],
the fused operation flag in bits [9:8], the PRP/SGL data transfer type in
bits [15:14], and the Command ID (CID) in bits [31:16]. The CID is used to
correlate completions with submissions and must be unique among outstanding
commands in a given SQ.

\begin{table}[htbp]
\centering
\caption{NVMe Submission Queue Entry (64 bytes)}
\label{tab:sqe-format}
\footnotesize
\begin{tabular}{@{}llp{4.0cm}@{}}
\toprule
\textbf{Offset} & \textbf{Field} & \textbf{Description} \\
\midrule
\texttt{00--03} & CDW0  & Opcode [7:0], FUSE [9:8], PSDT [15:14], CID [31:16] \\
\texttt{04--07} & NSID  & Namespace Identifier \\
\texttt{08--0F} & Rsvd  & Reserved (8 bytes) \\
\texttt{10--13} & MPTR  & Metadata Pointer (low 32) \\
\texttt{14--17} & MPTR  & Metadata Pointer (high 32) \\
\texttt{18--1B} & PRP1  & PRP Entry 1 (low 32) \\
\texttt{1C--1F} & PRP1  & PRP Entry 1 (high 32) \\
\texttt{20--23} & PRP2  & PRP Entry 2 (low 32) \\
\texttt{24--27} & PRP2  & PRP Entry 2 (high 32) \\
\texttt{28--2B} & CDW10 & Command-specific DWORD 10 \\
\texttt{2C--2F} & CDW11 & Command-specific DWORD 11 \\
\texttt{30--33} & CDW12 & Command-specific DWORD 12 \\
\texttt{34--37} & CDW13 & Command-specific DWORD 13 \\
\texttt{38--3B} & CDW14 & Command-specific DWORD 14 \\
\texttt{3C--3F} & CDW15 & Command-specific DWORD 15 \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Doorbell Mechanism}

The host writes command entries to the SQ in host memory, then notifies the
controller by writing the new tail pointer to the SQ Tail Doorbell register.
Doorbell registers are memory-mapped at BAR0 offsets starting at
\texttt{0x1000}: the Admin SQ Tail Doorbell is at BAR0+\texttt{0x1000}, the
Admin CQ Head Doorbell is at BAR0+\texttt{0x1004}, and I/O queue doorbells
follow at subsequent offsets defined by the doorbell stride
(\texttt{CAP.DSTRD}).

The Completion Queue Entry (CQE) is 16 bytes and includes a Status Field
(SCT and SC) that indicates command success or failure, the SQ Head pointer
(SQHD), the SQ Identifier (SQID), the Command ID (CID) for correlation with
submissions, and a Phase Tag (P) that toggles each time the CQ wraps to
distinguish new entries from stale ones.

\subsubsection{Relevant Admin Commands}

For recovery purposes, the relevant NVMe admin commands are:

\begin{description}[nosep,style=unboxed,leftmargin=0.5cm]
  \item[Format NVM (0x80)] Reformats the media, optionally performing
    secure erase. CDW10 contains LBAF (LBA format, bits [3:0]) and SES
    (Secure Erase Settings, bits [11:9]).
  \item[Sanitize (0x84)] More thorough data removal than Format. CDW10
    contains SANACT (Sanitize Action, bits [2:0]): 1=Exit Failure Mode,
    2=Block Erase, 3=Overwrite, 4=Crypto Erase. Clears all caches and
    over-provisioned areas.
  \item[Set Features (0x09)] Modifies controller features. CDW10 contains
    FID (Feature Identifier, bits [7:0]) and SV (Save, bit 31). CDW11
    contains the feature-specific value.
\end{description}

\subsection{USB-NVMe Bridges}
\label{sec:usb-bridges}

USB-to-NVMe bridge chips translate between the SCSI command set used by the
USB Mass Storage (Bulk-Only Transport, BOT) and UAS (USB Attached SCSI)
protocols and the NVMe command set. Common bridge chipsets include the
JMicron JMS583, Realtek RTL9210, and ASMedia ASM2362. Each implements its
own proprietary mechanism for passing NVMe admin commands from the host.

\subsubsection{ASM2362 Architecture}

The ASMedia ASM2362 is built around an 8051-compatible microcontroller core
running at approximately 114.3 MHz~\cite{cyrozap-usb-pcie}. Key architectural
features include:

\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{XRAM}: 64KB of mapped internal memory, accessible via the
    8051's MOVX instructions. Portions of XRAM are DMA-mapped to serve as
    NVMe queue structures, enabling the bridge to mediate between USB and
    PCIe without host memory involvement.
  \item \textbf{PCIe interface}: Gen 3 x2 lanes connecting to the NVMe
    device, with a TLP (Transaction Layer Packet) generation engine
    accessible from XRAM registers at \texttt{0xB200}--\texttt{0xB296}.
  \item \textbf{SPI flash}: External flash storage for firmware. Notably,
    \textbf{no firmware signature verification} is performed on updates,
    meaning arbitrary code can be loaded via the 0xE3 command.
  \item \textbf{UART debug}: 921600 baud 8N1 at 3.3V on pins 62/63,
    providing debug console access to the 8051 firmware.
  \item \textbf{USB ID}: VID \texttt{0x174C}, PID \texttt{0x2362}.
\end{itemize}

\subsubsection{The 0xE6 Passthrough CDB}
\label{sec:e6-passthrough}

The ASM2362 implements a vendor-specific 16-byte SCSI CDB with opcode
\texttt{0xE6} for tunneling NVMe admin commands over USB. This is the
mechanism used by smartmontools~\cite{smartmontools} (via the
\texttt{sntasmedia\_device} class) and Silicon Power's SP Toolbox
application. Table~\ref{tab:e6-cdb} shows the CDB byte layout.

\begin{table}[htbp]
\centering
\caption{ASM2362 0xE6 Passthrough CDB Format (16 bytes)}
\label{tab:e6-cdb}
\footnotesize
\begin{tabular}{@{}clp{3.8cm}@{}}
\toprule
\textbf{Byte} & \textbf{Field} & \textbf{Description} \\
\midrule
0     & \texttt{0xE6}      & ASMedia passthrough opcode \\
1     & NVMe OPC           & NVMe admin opcode \\
2     & Reserved           & \texttt{0x00} \\
3     & CDW10[7:0]         & Low byte of CDW10 \\
4--5  & Reserved           & \texttt{0x00 0x00} \\
6     & CDW10[23:16]       & Bits 23--16 of CDW10 \\
7     & CDW10[31:24]       & High byte of CDW10 \\
8--11 & CDW13              & Big-endian \\
12--15& CDW12              & Big-endian \\
\bottomrule
\end{tabular}
\end{table}

Two critical limitations exist with this CDB format. First, CDW10 bits
[15:8] and CDW11, CDW14, CDW15 cannot be transmitted, preventing full use
of commands that require these fields (e.g., the SEL field for Get Features
lives in CDW10[10:8] and is silently lost). Second, and more importantly for
recovery, the bridge firmware implements an \textbf{opcode whitelist} that
only forwards two NVMe admin opcodes:

\begin{itemize}[leftmargin=*,nosep]
  \item \texttt{0x02} -- Get Log Page (SMART, Error Log, etc.)
  \item \texttt{0x06} -- Identify (Controller and Namespace)
\end{itemize}

All other opcodes---including Format NVM (\texttt{0x80}), Sanitize
(\texttt{0x84}), Set Features (\texttt{0x09}), Get Features (\texttt{0x0A}),
Security Send (\texttt{0x81}), and Security Receive (\texttt{0x82})---are
silently dropped at the bridge firmware level. The SCSI layer returns
\texttt{GOOD} status to the host, but the command never reaches the NVMe
controller. This ``silent drop'' behavior is particularly insidious for
debugging, as it creates the appearance that the NVMe controller rejected
or ignored the command.

\subsubsection{SCSI-to-NVMe Translation}

The bridge's SCSI-to-NVMe Translation Layer (SNTL) also translates standard
SCSI commands to their NVMe equivalents for basic block I/O:

\begin{itemize}[leftmargin=*,nosep]
  \item SCSI READ(10/16) $\to$ NVMe Read
  \item SCSI WRITE(10/16) $\to$ NVMe Write
  \item SCSI INQUIRY $\to$ NVMe Identify (portions)
  \item SCSI MODE SENSE $\to$ synthesized from NVMe data
  \item SCSI READ CAPACITY $\to$ NVMe Identify Namespace
\end{itemize}

Importantly, the ``Medium not present'' error (Sense Key \texttt{0x02},
ASC \texttt{0x3A}) returned for admin commands is a \textbf{SCSI concept
that does not exist in NVMe}. The bridge generates this error when the NVMe
controller fails to respond as expected during initialization, not when an
NVMe command explicitly fails.

\subsection{Phison PS5012-E12 Controller}
\label{sec:phison}

The Phison PS5012-E12 is a quad-core, eight-channel NVMe 1.3 controller
widely used in consumer SSDs including Silicon Power, Kingston A2000,
Corsair MP510, and Sabrent Rocket products. It supports PCIe Gen 3 x4
and drives up to 2TB with 3D TLC NAND. Key characteristics of the specific
drive under investigation:

\begin{table}[htbp]
\centering
\caption{NVMe controller identification}
\label{tab:controller-id}
\footnotesize
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Vendor ID (VID) & \texttt{0x1E4B} (Phison Electronics) \\
Model Number & SPCC M.2 PCIe SSD \\
Serial Number & 00000000000000001387 \\
Firmware Revision & H211011a \\
OACS & \texttt{0x0006} (Format NVM + FW Download) \\
SANICAP & \texttt{0x00000002} (Block Erase only) \\
Capacity & 256 GB (500,118,192 $\times$ 512B) \\
\bottomrule
\end{tabular}
\end{table}

The OACS (Optional Admin Command Support) value of \texttt{0x0006}
indicates that the controller supports Format NVM (bit 1) and Firmware
Download/Commit (bit 2), but does not advertise Security Send/Receive
(bit 0) or Namespace Management (bit 3). The SANICAP (Sanitize Capabilities)
value of \texttt{0x00000002} indicates support for Block Erase (bit 1) only;
Crypto Erase (bit 0) and Overwrite (bit 2) are not supported.

\subsection{Related Work}
\label{sec:related}

The foundational reverse engineering of ASMedia USB bridge chips was
performed by cyrozap~\cite{cyrozap-usb-pcie}, who documented the XRAM memory
map, vendor SCSI commands (0xE0--0xE8), PCIe TLP generation mechanism, and
the absence of firmware signature verification. This work provided the
essential memory layout information that made our injection technique
possible. The smartmontools project~\cite{smartmontools} implements the 0xE6
passthrough for SMART monitoring through its \texttt{sntasmedia\_device}
class, confirming the whitelist limitation by only implementing Identify and
Get Log Page. The ASMTool project~\cite{asmtool} provides firmware dumping
and analysis capabilities for ASMedia bridges.

Phison-specific recovery tools exist through the usbdev.ru
community~\cite{phison-reinitial}, including the PS5012 Reinitial Tool
(ECFM22.6) designed specifically for drives exhibiting SMART attribute
\#0 = 8 (read-only mode). These tools require Windows, vendor-specific
firmware matching via Flash ID2 probing, and the drive must be ``under
firmware'' (not in ROM mode). Professional recovery via PC-3000
SSD~\cite{acelab-phison,rossmann-ssd} is available but expensive
(\$300--\$1,500) and requires physical access to the bare drive.

The Linux \texttt{sg3\_utils}~\cite{sg3-utils} package provides user-space
SCSI passthrough capabilities via the \texttt{sg\_raw} command, which we
used for initial prototyping before developing our Zig-based tool. The
\texttt{nvme-cli}~\cite{nvme-cli} tool provides comprehensive NVMe
management through the native Linux NVMe driver, but cannot operate through
USB bridges. The \texttt{sedutil} project~\cite{sedutil} implements TCG Opal
self-encrypting drive management, which is a separate security mechanism
from the FTL write protection we encountered.

Prior work on SSD power fault resilience~\cite{zheng2013understanding} has
shown that 13 out of 15 consumer SSDs exhibit data loss during power fault
testing, establishing the prevalence of the failure mode that initiated our
investigation. Cai et al.~\cite{cai2017errors} provide a comprehensive
analysis of flash memory error mechanisms, including the FTL corruption
scenarios that lead to controller protection modes.

% ══════════════════════════════════════════════════════════════════════
\section{Problem Analysis}
\label{sec:problem}

\subsection{Failure Scenario}

The target drive---a Silicon Power SPCC M.2 PCIe SSD (256GB) with Phison
PS5012-E12 controller---experienced a combined Windows hibernate and
unexpected power loss event in January 2026. The drive was connected via an
ASMedia ASM2362 USB enclosure to a Linux host (Rocky Linux 10). After the
event, the drive exhibited a distinctive failure pattern: reads succeed but
all writes silently fail.

\begin{itemize}[leftmargin=*,nosep]
  \item SCSI READ commands succeed (all existing data readable)
  \item SCSI WRITE commands report success but data is unchanged on readback
  \item NVMe admin commands via 0xE6 return ``Medium not present''
    (Sense Key \texttt{0x02}, ASC \texttt{0x3A})
  \item No operating system write-protect flags are set
    (\texttt{blockdev --getro} returns 0, SCSI WP mode page shows WP=0)
\end{itemize}

\subsection{Silent Write Failure Characterization}

The drive's silent write failure was confirmed through three independent
tests, each demonstrating that write operations appear to succeed at the
SCSI protocol level but produce no change on the media.

\subsubsection{Test 1: Direct Sector Write}

\begin{lstlisting}[style=shell,caption={dd to sector 0: write reports success}]
$ sudo dd if=/dev/zero of=/dev/sdb \
    bs=512 count=1 conv=fsync oflag=direct
512 bytes copied, 0.000313659 s, 1.6 MB/s

$ sudo xxd -l 16 /dev/sdb
00000000: 33c0 8ed0 bc00 7c8e ...
# MBR boot code remains -- NOT zeroed
\end{lstlisting}

The \texttt{dd} command reported successful completion with a data rate of
1.6 MB/s, and the \texttt{fsync} and \texttt{oflag=direct} flags ensured
the write bypassed the page cache. Yet the MBR boot code at sector 0 was
unchanged.

\subsubsection{Test 2: Filesystem Signature Erasure}

\begin{lstlisting}[style=shell,caption={wipefs: reported erased, data unchanged}]
$ sudo wipefs --all --force /dev/sdb
/dev/sdb: 8 bytes erased at offset 0x200 (gpt)
/dev/sdb: 2 bytes erased at offset 0x1fe (PMBR)

$ sudo xxd -s 512 -l 8 /dev/sdb
00000200: 4546 4920 5041 5254
# "EFI PART" GPT header still present
\end{lstlisting}

\subsubsection{Test 3: GPT Partition Entry Write}

\begin{lstlisting}[style=shell,caption={Sector 2 write: byte-for-byte identical}]
BEFORE:
00000430: ... 4d00 6900 6300 7200  M.i.c.r.

$ sudo dd if=/dev/zero of=/dev/sdb \
    bs=512 count=1 seek=2 conv=fsync oflag=direct
512 bytes copied

AFTER:
00000430: ... 4d00 6900 6300 7200  M.i.c.r.
# Data byte-for-byte identical
\end{lstlisting}

Standard Linux disk tools---\texttt{fdisk}, \texttt{gdisk},
\texttt{blkdiscard}, \texttt{sg\_format}---all reported success but left the
drive data unchanged. The SCSI Write Protect mode page correctly reported
\texttt{WP=0}, and Linux's \texttt{blockdev --getro} returned 0,
confirming that no host-side write protection was active. The controller
was accepting and acknowledging write commands at the SCSI protocol level
but not committing them to NAND.

\subsection{SMART Analysis}

The SMART/Health log, retrieved through the 0xE6 passthrough (opcode
\texttt{0x02}, which is whitelisted), revealed the drive's internal state:

\begin{table}[htbp]
\centering
\caption{SMART/Health log data (pre-recovery)}
\label{tab:smart-pre}
\footnotesize
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Critical Warning            & \texttt{0x00} \\
Temperature                 & 32$^\circ$C \\
Available Spare             & 1\% \\
Available Spare Threshold   & 10\% \\
Percentage Used             & 2\% \\
Data Units Read             & 48,115,774 \\
Data Units Written          & 31,289,178 \\
Host Read Commands          & 601,476,827 \\
Host Write Commands         & 1,076,834,710 \\
Media and Data Integrity Errors & 292 \\
Unsafe Shutdowns            & 270 \\
Power On Hours              & 5,936 \\
Power Cycles                & 1,253 \\
\bottomrule
\end{tabular}
\end{table}

Several findings are notable:

\paragraph{SMART Critical Warning = 0x00.} This was the most surprising
result. Bit 3 of the Critical Warning byte (mask \texttt{0x08}) indicates
``media placed in read-only mode'' per the NVMe specification. Our initial
hypothesis was that this bit would be set, driving the write protection.
Instead, it was \texttt{0x00}---\textbf{no warning bits were set}. The
controller had entered read-only mode through an internal mechanism not
reflected in the standard SMART critical warning byte.

\paragraph{Available Spare = 1\%.} The drive's spare block pool was nearly
depleted (threshold is 10\%), indicating significant NAND wear. Combined
with 292 media integrity errors, this suggests the FTL corruption was
exacerbated by limited spare capacity for bad block remapping.

\paragraph{270 Unsafe Shutdowns.} The high count of unsafe shutdowns
(power losses during operation) indicates a history of conditions favorable
to FTL corruption.

\subsection{Why the SMART Bit 3 Hypothesis Was Wrong}
\label{sec:bit3-wrong}

The NVMe specification states that Critical Warning bit 3 ``indicates that
the media has been placed in read only mode.'' This led to an initial
hypothesis that clearing this bit (or the condition it reflects) would
restore write access. However, this hypothesis was flawed for three reasons:

\begin{enumerate}[leftmargin=*]
  \item Bit 3 was not set---the controller entered read-only mode without
    setting this flag, suggesting a firmware-internal protection path
    distinct from the SMART-reported state.
  \item Bit 3, when set, reflects autonomous firmware state persisted in the
    controller's NAND service area. It is not a register the host can write
    to; no standard NVMe Set Features command can clear
    it~\cite{disktuna-wp}.
  \item Feature 0x84 (Namespace Write Protection) is a completely separate
    mechanism from firmware-triggered read-only mode. The former is a
    host-initiated policy; the latter is a controller-initiated protection.
\end{enumerate}

\subsection{The 0xE6 Whitelist Discovery}

Through systematic testing of all NVMe admin opcodes observed in SP Toolbox
USB captures (obtained via Frida hooks on \texttt{DeviceIoControl} calls),
we confirmed the bridge firmware whitelist:

\begin{table}[htbp]
\centering
\caption{ASM2362 0xE6 opcode whitelist testing results}
\label{tab:whitelist}
\footnotesize
\begin{tabular}{@{}clcc@{}}
\toprule
\textbf{Opcode} & \textbf{Command} & \textbf{Passes?} & \textbf{Count} \\
\midrule
\texttt{0x02} & Get Log Page     & \color{successgreen}Yes & 17 \\
\texttt{0x06} & Identify         & \color{successgreen}Yes & 8 \\
\texttt{0x09} & Set Features     & \color{failred}No  & 6 \\
\texttt{0x0A} & Get Features     & \color{failred}No  & 9 \\
\texttt{0x80} & Format NVM       & \color{failred}No  & 5 \\
\texttt{0x81} & Security Send    & \color{failred}No  & 4 \\
\texttt{0x82} & Security Receive & \color{failred}No  & 14 \\
\texttt{0x84} & Sanitize         & \color{failred}No  & 6 \\
\bottomrule
\end{tabular}
\end{table}

The ``Count'' column shows the number of occurrences of each opcode in the
captured SP Toolbox session. The tool itself attempts all these commands
through the 0xE6 passthrough, apparently unaware (or unconcerned) that
most are silently dropped. This suggests that SP Toolbox may have been
designed for bridges with more permissive whitelists, or that it includes
fallback paths we did not capture.

% ══════════════════════════════════════════════════════════════════════
\section{XRAM Injection Methodology}
\label{sec:methodology}

\subsection{ASM2362 Vendor SCSI Commands}

Beyond the 0xE6 NVMe passthrough, the ASM2362 exposes additional
vendor-specific SCSI commands~\cite{cyrozap-usb-pcie}. Table~\ref{tab:vendor-cmds}
lists the complete set discovered through firmware reverse engineering.

\begin{table}[htbp]
\centering
\caption{ASM2362 vendor-specific SCSI commands}
\label{tab:vendor-cmds}
\footnotesize
\begin{tabular}{@{}cllc@{}}
\toprule
\textbf{Opcode} & \textbf{Command} & \textbf{Direction} & \textbf{Size} \\
\midrule
\texttt{0xE0} & Read Config  & Dev$\to$Host & 128B \\
\texttt{0xE1} & Write Config & Host$\to$Dev & 128B \\
\texttt{0xE2} & Flash Read   & Dev$\to$Host & Variable \\
\texttt{0xE3} & Firmware Write & Host$\to$Dev & Variable \\
\texttt{0xE4} & XDATA Read   & Dev$\to$Host & 1--255B \\
\texttt{0xE5} & XDATA Write  & None         & 1B (in CDB) \\
\texttt{0xE6} & NVMe Admin   & Dev$\to$Host & Variable \\
\texttt{0xE8} & Reset        & None         & N/A \\
\bottomrule
\end{tabular}
\end{table}

The 0xE4 and 0xE5 commands are the key enablers for our technique. They
provide direct read/write access to the bridge chip's 64KB XRAM address
space. Critically, the NVMe Admin Submission Queue is DMA-mapped into XRAM,
meaning that XRAM writes via 0xE5 are reflected in the queue structures
that the NVMe controller reads when processing commands.

\subsection{UAS vs.\ BOT Mode Requirement}
\label{sec:uas-bot}

An essential prerequisite discovered during implementation is that the vendor
SCSI commands \texttt{0xE4}, \texttt{0xE5}, and \texttt{0xE8}
\textbf{require the \texttt{usb-storage} driver (Bulk-Only Transport, BOT
mode)}, not the \texttt{uas} driver (USB Attached SCSI). Under UAS, these
commands consistently return \texttt{DID\_ERROR} with
\texttt{host\_status=0x07} and errno \texttt{-75} (\texttt{EOVERFLOW}).

This is likely because UAS implements its own command queuing and protocol
framing that does not expect the vendor-specific CDB formats. BOT mode
passes CDBs more transparently through the USB mass storage class driver.

\begin{lstlisting}[style=shell,caption={Switching ASM2362 from UAS to BOT mode}]
sudo bash -c '
  # Unbind the USB device
  echo "1-3" > /sys/bus/usb/drivers/usb/unbind
  # Remove both drivers
  rmmod uas
  rmmod usb_storage
  # Reload usb-storage with quirk flag
  modprobe usb-storage quirks=174c:2362:u
  # Rebind the device
  echo "1-3" > /sys/bus/usb/drivers/usb/bind
'
# Device re-enumerates as /dev/sdc
\end{lstlisting}

Verification that BOT mode is active:
\begin{lstlisting}[style=shell,caption={Verifying BOT mode}]
$ readlink /sys/devices/.../1-3:1.0/driver
# Must show: usb-storage (not uas)
\end{lstlisting}

The USB port number (\texttt{1-3} in this example) varies by system and
must be determined from \texttt{lsusb -t} or \texttt{dmesg} output.

\subsection{XRAM Memory Map}
\label{sec:xram-map}

Table~\ref{tab:xram-map} shows the XRAM memory regions relevant to NVMe
command injection, as documented by cyrozap~\cite{cyrozap-usb-pcie} and
refined through our own systematic probing via 0xE4 reads across the entire
64KB address space.

\begin{table}[htbp]
\centering
\caption{ASM2362 XRAM memory map (relevant regions)}
\label{tab:xram-map}
\footnotesize
\begin{tabular}{@{}llp{3.0cm}@{}}
\toprule
\textbf{Address} & \textbf{Size} & \textbf{Contents} \\
\midrule
\texttt{0x07F0}      & 6B    & Firmware version string \\
\texttt{0x9000+}     & Var   & USB/SCSI control registers \\
\texttt{0xA000--AFFF} & 4KB   & NVMe I/O Submission Queue \\
\texttt{0xB000--B0FF} & 256B  & Admin SQ (4 entries $\times$ 64B) \\
\texttt{0xB100--B1FF} & 256B  & Admin SQ (unmapped/unused) \\
\texttt{0xB200--B209} & 10B   & PCIe config space buffer \\
\texttt{0xB210--B21B} & 12B   & TLP request header (3$\times$u32 BE) \\
\texttt{0xB220--B223} & 4B    & TLP write data (u32 BE) \\
\texttt{0xB224--B22F} & 12B   & TLP completion header \\
\texttt{0xB254}       & 1B    & TLP operation trigger \\
\texttt{0xB284}       & 1B    & Config request validation \\
\texttt{0xB296}       & 1B    & TLP control/status register \\
\texttt{0xB800--B9FF} & 512B  & I/O Completion Queue (SQID=1) \\
\texttt{0xBC00--BC7F} & 128B  & Admin CQ (8 $\times$ 16B) \\
\texttt{0xF000--FFFF} & 4KB   & NVMe data buffer \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Admin SQ Depth Discovery}

A significant discovery was that the \textbf{Admin SQ depth is 4, not 8}.
Although the XRAM region \texttt{0xB000--0xB1FF} provides 512 bytes (space
for 8 entries of 64 bytes each), only slots 0--3 are within the queue
boundary as configured by the bridge firmware. This was confirmed through
two observations:

\begin{enumerate}[leftmargin=*]
  \item \textbf{SQHD wrapping}: Completion Queue entries showed the SQ Head
    pointer (SQHD) wrapping at 4, never reaching values of 4--7.
  \item \textbf{Slot 4--7 injection}: Commands written to XRAM addresses
    \texttt{0xB100}--\texttt{0xB1FF} (slots 4--7) were never consumed by
    the controller---no corresponding CQ entries appeared.
\end{enumerate}

The four slots at boot time contain the bridge firmware's NVMe initialization
sequence: Set Features (Number of Queues), Set Features (other), Create I/O
CQ, Create I/O SQ. All four are occupied, meaning injection must overwrite
an existing entry.

\subsubsection{Admin CQ Base Discovery}

Similarly, the Admin CQ base address was initially assumed to be
\texttt{0xB800} based on the documentation, but systematic probing revealed
that \texttt{0xB800} contains I/O Completion Queue entries (identified by
SQID=1 in the entries). The actual Admin CQ (SQID=0) is at
\texttt{0xBC00}, containing 8 entries of 16 bytes each (128 bytes total).

\subsection{CDB Formats for XRAM Access}
\label{sec:cdb-formats}

The vendor SCSI CDB formats were corrected from our implementation
experience. The original cyrozap documentation lists the byte positions
correctly, but early implementations in our codebase had the address and
value bytes transposed, leading to writes to incorrect XRAM addresses.

\vspace{0.5em}
\noindent\textbf{0xE4 -- XDATA Read (6-byte CDB):}
\begin{lstlisting}[style=hex]
Byte 0: E4           ; Opcode
Byte 1: <length>     ; Bytes to read (1-255)
Byte 2: 00           ; Padding
Byte 3: <addr_hi>    ; Address bits [15:8]
Byte 4: <addr_lo>    ; Address bits [7:0]
Byte 5: 00           ; Padding
Direction: SG_DXFER_FROM_DEV
Transfer length: <length> bytes
\end{lstlisting}

\noindent\textbf{0xE5 -- XDATA Write (6-byte CDB):}
\begin{lstlisting}[style=hex]
Byte 0: E5           ; Opcode
Byte 1: <value>      ; Single byte to write
Byte 2: 00           ; Padding
Byte 3: <addr_hi>    ; Address bits [15:8]
Byte 4: <addr_lo>    ; Address bits [7:0]
Byte 5: 00           ; Padding
Direction: SG_DXFER_NONE
(Byte is embedded in CDB, no data transfer)
\end{lstlisting}

\noindent\textbf{0xE8 -- Reset (12-byte CDB):}
\begin{lstlisting}[style=hex]
Byte 0:  E8          ; Opcode
Byte 1:  <type>      ; 0x00=CPU, 0x01=PCIe
Bytes 2-11: 00       ; Padding
Direction: SG_DXFER_NONE
\end{lstlisting}

The 0xE5 write command embeds the data byte \emph{within the CDB
itself}---there is no SCSI data transfer phase. This means writing a 64-byte
NVMe SQE requires issuing 64 separate 0xE5 SCSI commands, each writing one
byte. While slow (approximately 2--3 seconds for a complete SQE), this is
reliable and provides byte-level write verification capability: after each
0xE5 write, an 0xE4 read of the same address can confirm the value was
stored correctly.

The maximum read length for 0xE4 is 255 bytes per command. Reading the
full 256-byte Admin SQ region requires two reads:
\texttt{xdataRead(0xB000, 255)} and \texttt{xdataRead(0xB0FF, 1)},
or equivalent chunked reads.

\subsection{PCIe TLP Doorbell Mechanism}
\label{sec:pcie-tlp}

After writing the NVMe command into the Admin SQ, the SQ Tail Doorbell must
be rung to notify the NVMe controller of the new submission. The ASM2362's
PCIe TLP generation engine allows crafting arbitrary PCIe Transaction Layer
Packets from XRAM registers. The procedure, ported from
cyrozap's \texttt{pcie\_gen\_req()} Python function~\cite{cyrozap-usb-pcie}:

\begin{enumerate}[leftmargin=*]
  \item \textbf{Read BAR0} from PCIe config space via a Type 1 Configuration
    Read TLP. The address encodes bus=1, device=0, function=0,
    register=\texttt{0x10}: \\
    \texttt{cfg\_addr = (1 << 24) | (0 << 19) | (0 << 16) | 0x10}. \\
    Result for our device: BAR0 = \texttt{0x00D00000} (masked to clear
    type bits).

  \item \textbf{Calculate doorbell address}: Admin SQ Tail Doorbell =
    BAR0 + \texttt{0x1000} = \texttt{0x00D01000}.

  \item \textbf{Write new tail value} (big-endian u32) to the TLP data
    register at XRAM \texttt{0xB220}. For tail=1:
    \texttt{0x00000001} $\to$ bytes \texttt{00 00 00 01}.

  \item \textbf{Write 12-byte TLP header} (3 $\times$ u32, big-endian)
    to XRAM \texttt{0xB210}:
    \begin{itemize}[nosep]
      \item DW0: \texttt{0x40000001} -- fmt\_type=\texttt{0x40} (Memory
        Write, 3DW header, with data), length=1 DW
      \item DW1: byte enable mask (first byte enable field)
      \item DW2: \texttt{0x00D01000} -- masked doorbell address
    \end{itemize}

  \item \textbf{Clear timeout bit}: write \texttt{0x01} to XRAM
    \texttt{0xB296} (CSR).

  \item \textbf{Trigger TLP engine}: write \texttt{0x0F} to XRAM
    \texttt{0xB254} (operation register).

  \item \textbf{Poll CSR for ready}: read XRAM \texttt{0xB296} until
    bit 2 is set (up to 1000 iterations).

  \item \textbf{Send TLP}: write \texttt{0x04} to XRAM \texttt{0xB296}.
\end{enumerate}

This is a \emph{posted} memory write TLP (fmt\_type \texttt{0x40}),
meaning no completion TLP is expected from the NVMe controller. The USB
connection stays alive throughout the entire procedure---unlike using the
0xE8 PCIe reset command (type=1), which resets the PCIe link and causes
USB disconnection and device re-enumeration.

\paragraph{Limitation: No PCIe memory reads.} While the TLP engine supports
PCIe configuration reads and posted memory writes, non-posted memory
\emph{read} TLPs hang or timeout. This means we cannot directly read the
NVMe controller's MMIO registers (CSTS, CC, AQA, etc.) or completion queue
entries in controller memory. We can only read queue structures that the
bridge firmware has mapped into XRAM.

% ══════════════════════════════════════════════════════════════════════
\section{Failed Approaches}
\label{sec:failures}

Before arriving at the successful injection technique, numerous approaches
were attempted over a period of approximately five weeks (late January
through early March 2026). Table~\ref{tab:failures} catalogs these dead
ends chronologically.

\begin{table*}[htbp]
\centering
\caption{Chronological summary of failed recovery approaches}
\label{tab:failures}
\footnotesize
\begin{tabularx}{\textwidth}{@{}llXX@{}}
\toprule
\textbf{Date} & \textbf{Approach} & \textbf{Method} & \textbf{Failure Reason} \\
\midrule
Jan 21 & Standard tools &
  dd, wipefs, fdisk, gdisk, blkdiscard, sg\_format &
  All report success but produce no change: firmware-level silent write failure \\
\addlinespace
Jan 25 & 0xE6 Format NVM &
  Build 0xE6 CDB with opcode \texttt{0x80}, various LBAF/SES values &
  Bridge firmware whitelist silently drops; SCSI returns GOOD but NVMe controller never receives command \\
\addlinespace
Jan 26 & 0xE6 Sanitize &
  Build 0xE6 CDB with opcode \texttt{0x84}, SANACT=2 and SANACT=4 &
  Same whitelist: silently dropped \\
\addlinespace
Jan 28 & 0xE6 Security Send &
  ATA password protocol (\texttt{0xEF}) via opcode \texttt{0x81} to clear security state &
  Whitelist blocks opcodes \texttt{0x81}/\texttt{0x82} \\
\addlinespace
Jan 30 & 0xE6 Set Features &
  FID \texttt{0x84} (Namespace Write Protect) to disable write protection &
  Whitelist blocks opcode \texttt{0x09}; also, NWPC is a separate mechanism from FTL read-only \\
\addlinespace
Feb 1 & Wine + SP Toolbox &
  Run Silicon Power's Windows tool under Wine on Linux &
  Wine does not implement \texttt{IOCTL\_SCSI\_PASS\_THROUGH\_DIRECT} (\texttt{0x4D014}) \\
\addlinespace
Feb 15 & SMART bit 3 clearing &
  Hypothesis: clear Critical Warning bit 3 to restore writes &
  Bit 3 = 0x00 (not set); hypothesis was wrong \\
\addlinespace
Feb 28 & XRAM inject slots 4--7 &
  Write NVMe command to Admin SQ entries 4--7 via 0xE5 &
  Admin SQ depth = 4; entries beyond slot 3 are outside queue boundary \\
\addlinespace
Mar 1 & Sanitize Crypto Erase &
  Inject Sanitize with SANACT=4 (Crypto Erase) to slot 0 &
  SANICAP = \texttt{0x00000002}: Crypto Erase not supported \\
\addlinespace
Mar 2 & Format NVM via inject &
  Inject Format NVM (\texttt{0x80}) to slot 0, ring doorbell &
  Controller accepts (CQ success), but FTL write protection persists \\
\addlinespace
Mar 3 & PCIe reset doorbell &
  Use 0xE8 type=1 to trigger SQ processing &
  Disconnects USB; Format during reset caused 0B capacity (reverted on power cycle) \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{The Format NVM Paradox}

An instructive failure was the Format NVM command via XRAM injection
(March 2). The command was injected into Admin SQ slot 0, the PCIe TLP
doorbell was rung with tail=1, and the Completion Queue showed a success
entry with the correct CID, SCT=0, SC=0. Yet the drive remained in
read-only mode afterward.

This reveals an important distinction: on the Phison PS5012-E12, Format NVM
affects the logical namespace structure (LBA format, protection information)
but does \textbf{not} reset the FTL's internal protection state. Only
Sanitize, which performs a physical NAND block erase including
over-provisioned areas and internal metadata, was able to clear the FTL
corruption flag and restore write access.

This behavior may be controller-specific. Other NVMe controllers may respond
differently to Format NVM in similar scenarios. However, the general
principle holds: Sanitize is a more thorough operation that affects the
physical media, while Format is primarily a logical operation.

\subsection{The Queue Depth Discovery}

Injecting commands into slots 4--7 (February 28) produced no Completion
Queue entries and no observable effect. By dumping the Admin CQ at
\texttt{0xBC00} after injecting to various slots, we observed that the SQHD
(SQ Head pointer) in completion entries never exceeded 3, confirming a queue
depth of 4. Additionally, the XRAM region at \texttt{0xB100--0xB1FF} (where
slots 4--7 would reside) contained zeros even after the bridge's
initialization sequence, whereas \texttt{0xB000--0xB0FF} contained the four
initialization commands.

\subsection{The PCIe Reset Disaster}

Using the 0xE8 PCIe reset (type=1) as a ``doorbell'' mechanism (March 3) had
two problems. First, it disconnects USB, requiring device re-enumeration.
Second, when combined with a pending Format NVM command, the reset caused the
drive to report 0 bytes capacity after re-enumeration. Fortunately, this was
reverted by a full power cycle (physically disconnecting the USB enclosure).
This experience led to the development of the proper PCIe TLP doorbell
mechanism, which writes to the doorbell register without disrupting the PCIe
link or USB connection.

% ══════════════════════════════════════════════════════════════════════
\section{Successful Recovery}
\label{sec:recovery}

\subsection{SANICAP Analysis}

Before attempting the Sanitize command, we verified the controller's
capabilities through the Identify Controller data (retrieved via the
whitelisted 0xE6 Identify command). The relevant fields:

\begin{lstlisting}[style=hex,caption={Controller capability analysis}]
OACS    = 0x0006
  Bit 0 (Security Send/Recv): 0  Not supported
  Bit 1 (Format NVM):         1  SUPPORTED
  Bit 2 (Firmware Download):  1  Supported

SANICAP = 0x00000002
  Bit 0 (Crypto Erase):       0  Not supported
  Bit 1 (Block Erase):        1  SUPPORTED
  Bit 2 (Overwrite):          0  Not supported
\end{lstlisting}

Only Sanitize Block Erase (SANACT=2) was available. Our earlier attempt with
Crypto Erase (SANACT=4) had failed specifically because it is not supported
by this controller. The OACS field also revealed that Format NVM \emph{is}
supported (bit 1), explaining why the controller accepted the Format command
even though it did not clear the FTL protection.

\subsection{Crafting the Sanitize SQE}

The Sanitize Block Erase command is encoded as a 64-byte NVMe SQE:

\begin{lstlisting}[style=hex,caption={Sanitize Block Erase SQE (little-endian)}]
Offset  Bytes                     Field
------  -----                     -----
00-03:  84 00 42 42               CDW0
        OPC=0x84 (Sanitize)
        CID=0x4242 (correlation)
04-07:  00 00 00 00               NSID=0 (global)
08-0F:  00 00 00 00 00 00 00 00   Reserved
10-17:  00 00 00 00 00 00 00 00   MPTR (unused)
18-1F:  00 00 00 00 00 00 00 00   PRP1 (unused)
20-27:  00 00 00 00 00 00 00 00   PRP2 (unused)
28-2B:  02 00 00 00               CDW10
        SANACT=2 (Block Erase)
2C-3F:  00 00 00 00 ... 00 00     CDW11-15 (zeros)
\end{lstlisting}

Key design choices:
\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{CID = 0x4242}: A distinctive value chosen for easy visual
    identification in CQ hex dumps.
  \item \textbf{NSID = 0}: Sanitize operates on the entire controller, not
    a specific namespace.
  \item \textbf{PRP1/PRP2 = 0}: Sanitize does not transfer data; these
    fields are unused.
  \item \textbf{CDW10 = 0x00000002}: SANACT=2 (Block Erase) in bits [2:0].
\end{itemize}

\subsection{Injection Procedure}

The successful recovery was performed on March 4, 2026. The complete
command sequence:

\subsubsection{Step 1: Switch to BOT Mode}

\begin{lstlisting}[style=shell]
sudo modprobe usb-storage quirks=174c:2362:u
# Verify: /dev/sda via usb-storage driver
\end{lstlisting}

\subsubsection{Step 2: Probe and Verify Bridge}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool probe /dev/sda
# Output: ASM2362 detected
# Firmware: 19 05 02 81 16 00
\end{lstlisting}

\subsubsection{Step 3: Identify Controller}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool identify /dev/sda
# VID: 0x1E4B (Phison)
# Model: SPCC M.2 PCIe SSD
# FW: H211011a
# OACS: 0x0006, SANICAP: 0x00000002
\end{lstlisting}

\subsubsection{Step 4: Inspect Admin SQ State}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool admin-cq /dev/sda
# Shows 4 occupied slots (bridge init sequence)
# Slots 0-3: Set Features, Set Features,
#            Create I/O CQ, Create I/O SQ
\end{lstlisting}

\subsubsection{Step 5: Dry-Run Injection}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool inject \
  --inject-cmd=sanitize-block \
  --slot=0 --tail=1 --cid=0x4242 \
  --dry-run /dev/sda
# Phase 1: Read Admin SQ ... OK
# Phase 2: Using slot 0 at 0xB000
# Phase 3: Writing 64 bytes ... OK
# Phase 4: Verification ... MATCH
# Phase 5: SKIPPED (dry-run)
\end{lstlisting}

\subsubsection{Step 6: Live Injection}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool inject \
  --inject-cmd=sanitize-block \
  --slot=0 --tail=1 --cid=0x4242 \
  --force /dev/sda
# Phase 1-4: (same as dry-run)
# Phase 5: Ringing doorbell, tail=1
#   BAR0 = 0x00D00000
#   Doorbell addr = 0x00D01000
#   Doorbell rung successfully (USB alive)
# Phase 6: Post-injection state read ... OK
\end{lstlisting}

\subsubsection{Step 7: Verify Completion}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool admin-cq /dev/sda
\end{lstlisting}

The Admin CQ at XRAM \texttt{0xBC00} showed a completion entry:

\begin{lstlisting}[style=hex,caption={Admin CQ completion entry}]
BC00: xx xx xx xx  ; Command-specific
BC04: 42 42 xx xx  ; CID=0x4242 (match!)
BC08: xx xx xx xx  ; SQHD, SQID
BC0C: 00 00 01 xx  ; SCT=0 SC=0 Phase=1
                    ; Status: SUCCESS
\end{lstlisting}

\subsubsection{Step 8: Bridge Reset}

\begin{lstlisting}[style=shell]
sudo ./zig-out/bin/asm2362-tool reset \
  --reset-type=0 /dev/sda
# CPU reset: 8051 restarts
# NVMe controller re-initialized from erased NAND
# Device re-enumerates on USB
\end{lstlisting}

\subsubsection{Step 9: Post-Recovery Verification}

\begin{lstlisting}[style=shell,caption={Write verification after recovery}]
# Write 1MB of random data
$ sudo dd if=/dev/urandom of=/dev/sda \
    bs=1M count=1 oflag=direct
1+0 records out
1048576 bytes copied, 17.7 MB/s

# Read back and verify
$ sudo dd if=/dev/sda bs=1M count=1 \
    | md5sum
# Hash matches -- writes functional!
\end{lstlisting}

\subsection{Post-Recovery State}

After the Sanitize Block Erase and bridge reset, the drive was fully
operational:

\begin{table}[htbp]
\centering
\caption{Post-recovery verification results}
\label{tab:post-recovery}
\footnotesize
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Check} & \textbf{Result} \\
\midrule
Sanitize Status (SSTAT)     & \texttt{0x0001} (Complete) \\
Write test (1MB random)     & Success, 17.7 MB/s \\
Read-back MD5 verification  & Match confirmed \\
Identify Controller         & Responds normally \\
SMART Critical Warning      & \texttt{0x00} \\
NVMe admin commands         & All functional \\
Drive capacity              & 256 GB (restored) \\
\bottomrule
\end{tabular}
\end{table}

All previous data was erased by the Sanitize Block Erase operation, as
expected. The drive was subsequently reformatted with a fresh filesystem and
returned to service.

% ══════════════════════════════════════════════════════════════════════
\section{Implementation}
\label{sec:implementation}

\subsection{Tool Architecture}

The recovery tool (\texttt{asm2362-tool}) is implemented in
Zig~\cite{zig-lang} (version 0.13.0+), chosen for its low-level control,
compile-time safety, and C interoperability (needed for Linux kernel
headers). The tool comprises approximately 4,000 lines of code with 35 unit
tests organized in a layered architecture.

\begin{lstlisting}[style=hex,caption={Source tree structure}]
src/
  main.zig            CLI entry point (900+ lines)
  log.zig             JSON command logging
  scsi/
    sg_io.zig         Linux SG_IO ioctl wrapper
    sense.zig         SCSI sense data parsing
  asm2362/
    passthrough.zig   0xE6 CDB builders (700 lines)
    commands.zig      NVMe command helpers
    xram.zig          0xE4/0xE5/0xE8, injection,
                      PCIe TLP doorbell (1100 lines)
  nvme/
    identify.zig      Identify Controller/NS parsing
  frida/
    hooks.js          Windows DeviceIoControl hooks
    replay.zig        Frida capture replay
  analysis/
    probe.zig         Bridge capability detection
\end{lstlisting}

\subsection{CLI Command Interface}

The tool provides six commands:

\begin{table}[htbp]
\centering
\caption{CLI command reference}
\label{tab:cli-cmds}
\footnotesize
\begin{tabular}{@{}lp{4.5cm}@{}}
\toprule
\textbf{Command} & \textbf{Purpose} \\
\midrule
\texttt{probe}    & Detect ASM2362 bridge, read firmware version via 0xE4 \\
\texttt{identify} & NVMe Identify Controller/Namespace via 0xE6 passthrough \\
\texttt{smart}    & Read SMART/Health log (supports \texttt{--json}) \\
\texttt{inject}   & Inject NVMe command into Admin SQ via XRAM bypass \\
\texttt{admin-cq} & Dump Admin Completion Queue entries from XRAM \\
\texttt{reset}    & Send bridge reset (CPU or PCIe) \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{inject} command accepts several parameters for precise control:

\begin{lstlisting}[style=shell,caption={Inject command parameters}]
--inject-cmd=CMD     # format, sanitize-block,
                     # sanitize-crypto, clear-wp
--slot=N             # Explicit SQ slot (0-3)
--tail=N             # Explicit doorbell tail value
--cid=0xNNNN         # Command ID for CQ correlation
--nsid=N             # Namespace ID (default 0xFFFFFFFF)
--force              # Required for live execution
--dry-run            # Write SQ but skip doorbell
\end{lstlisting}

\subsection{SG\_IO Interface Layer}

All SCSI commands are issued via the Linux \texttt{SG\_IO}
ioctl~\cite{linux-sg-io}, which provides synchronous SCSI command passthrough
to block devices. The \texttt{sg\_io.zig} module wraps the
\texttt{sg\_io\_hdr} structure and handles:

\begin{itemize}[leftmargin=*,nosep]
  \item CDB construction and submission
  \item Data buffer allocation and direction control
  \item Sense buffer parsing for CHECK CONDITION status
  \item Timeout management (configurable per-command)
  \item Error mapping from Linux errno to typed Zig errors
\end{itemize}

\subsection{Key Data Structures}

The \texttt{NvmeSqEntry} structure models the 64-byte NVMe SQE with
serialization to/from little-endian byte arrays:

\begin{lstlisting}[style=zig,caption={NvmeSqEntry structure (Zig)}]
pub const NvmeSqEntry = struct {
    cdw0: u32,     // OPC | FUSE | CID<<16
    nsid: u32,
    reserved8: u32 = 0,
    reserved12: u32 = 0,
    mptr_lo: u32 = 0,
    mptr_hi: u32 = 0,
    prp1_lo: u32 = 0,
    prp1_hi: u32 = 0,
    prp2_lo: u32 = 0,
    prp2_hi: u32 = 0,
    cdw10: u32 = 0,
    cdw11: u32 = 0,
    cdw12: u32 = 0,
    cdw13: u32 = 0,
    cdw14: u32 = 0,
    cdw15: u32 = 0,

    /// Serialize to 64 little-endian bytes
    pub fn toBytes(self: NvmeSqEntry) [64]u8;

    /// Deserialize from 64 LE XRAM bytes
    pub fn fromBytes(
        bytes: []const u8
    ) NvmeSqEntry;
};
\end{lstlisting}

Command crafting functions build SQEs for specific NVMe commands:

\begin{lstlisting}[style=zig,caption={Command crafting functions}]
pub fn craftSanitizeEntry(
    sanact: u3,
    command_id: u16,
) NvmeSqEntry {
    return .{
        .cdw0 = @as(u32, 0x84)
               | (@as(u32, command_id) << 16),
        .nsid = 0,
        .cdw10 = @as(u32, sanact),
    };
}

pub fn craftFormatNvmEntry(
    nsid: u32,
    lbaf: u4,
    ses: u3,
    command_id: u16,
) NvmeSqEntry {
    const cdw10 = @as(u32, lbaf)
                | (@as(u32, ses) << 9);
    return .{
        .cdw0 = @as(u32, 0x80)
               | (@as(u32, command_id) << 16),
        .nsid = nsid,
        .cdw10 = cdw10,
    };
}
\end{lstlisting}

\subsection{Injection Pipeline}

The \texttt{injectCommand()} function implements the six-phase injection
pipeline described in Section~\ref{sec:recovery}:

\begin{enumerate}[leftmargin=*,nosep]
  \item \textbf{Read} current Admin SQ state via 0xE4 (256 bytes from
    \texttt{0xB000}), parsing 4 $\times$ 64-byte entries.
  \item \textbf{Select} SQ slot: explicit (via \texttt{--slot}) or
    auto-detect first empty entry via \texttt{findEmptySlot()}.
  \item \textbf{Write} 64 bytes via 64 individual 0xE5 commands. Each byte
    is immediately verified via 0xE4 readback; any mismatch aborts the
    injection with \texttt{VerificationFailed}.
  \item \textbf{Verify} full 64-byte readback against expected SQE bytes
    using \texttt{std.mem.eql()}.
  \item \textbf{Doorbell}: Call \texttt{ringDoorbell()} with the specified
    tail value, which reads BAR0 via PCIe config TLP and writes the tail
    to BAR0+0x1000 via posted memory write TLP. Skipped in dry-run mode.
    Falls back to 0xE8 PCIe reset on doorbell failure.
  \item \textbf{Post-read} Admin SQ and CQ state for verification.
\end{enumerate}

The function returns an \texttt{InjectResult} struct containing the
pre/post SQ state, slot used, byte count, verification status, and
timing information, enabling the caller to present detailed diagnostics.

\subsection{Safety Mechanisms}

The tool implements multiple safety layers to prevent accidental damage:

\begin{description}[nosep,style=unboxed,leftmargin=0.5cm]
  \item[Dry-run default] All injection commands default to dry-run mode.
    The \texttt{--force} flag is required for live execution with doorbell.
  \item[Per-byte verification] Every XRAM byte write is verified via
    0xE4 readback before proceeding. A single mismatch aborts.
  \item[Address validation] Write operations are restricted to Admin SQ
    (\texttt{0xB000--B1FF}) and data buffer (\texttt{0xF000--FFFF}).
    Writes to MMIO registers (\texttt{0xB200+}) are blocked.
  \item[CID correlation] User-specified Command IDs (\texttt{--cid})
    enable unambiguous CQ entry correlation, preventing confusion with
    stale completion entries.
  \item[Explicit slot/tail] Queue slot and doorbell tail can be specified
    explicitly to avoid assumptions about queue state, which proved critical
    when the auto-detection logic made incorrect assumptions about queue
    depth.
\end{description}

\subsection{Testing}

The test suite includes 35 unit tests covering:

\begin{itemize}[leftmargin=*,nosep]
  \item CDB construction for all vendor commands (0xE4, 0xE5, 0xE8)
  \item CDB construction for all 0xE6 passthrough commands
  \item NvmeSqEntry serialization round-trips (toBytes/fromBytes)
  \item Command crafting (Format, Sanitize, Set Features)
  \item Queue slot finding (empty, partially full, completely full)
  \item Address safety validation (read vs.\ write regions)
  \item Little-endian helper functions
\end{itemize}

Tests run without hardware via \texttt{zig build test-all}. Hardware-dependent
testing requires root access and a connected ASM2362 device.

% ══════════════════════════════════════════════════════════════════════
\section{Discussion}
\label{sec:discussion}

\subsection{Implications for SSD Data Recovery}

The XRAM injection technique demonstrates that USB-to-NVMe bridges, despite
their firmware-level command restrictions, can be bypassed through direct
memory manipulation. This has significant implications:

\begin{itemize}[leftmargin=*]
  \item \textbf{USB recovery is possible}: Drives previously considered
    unrecoverable via USB may be salvageable without physical disassembly
    or M.2 slot access. This is particularly relevant for laptops and
    systems without available M.2 slots.
  \item \textbf{Non-destructive to bridge}: The technique only modifies
    volatile XRAM, which is reset on power cycle or CPU reset. No
    permanent changes are made to the bridge firmware or configuration.
  \item \textbf{Cost savings}: Professional recovery services using PC-3000
    SSD~\cite{acelab-phison,rossmann-ssd} cost \$300--\$1,500. Our technique
    requires only the USB enclosure already in use and an open-source tool.
  \item \textbf{Sanitize vs.\ Format}: The finding that Sanitize Block
    Erase (not Format NVM) is required to clear FTL write protection on
    the PS5012-E12 is valuable for the recovery community, as most
    guides recommend Format as the first attempt.
\end{itemize}

\subsection{Generalizability}

\subsubsection{Other ASMedia Bridges}

The technique should generalize to other ASMedia bridge variants in the
ASM236X family (ASM2362, ASM2364, etc.) that share the same 8051
architecture and vendor SCSI command set. The XRAM memory map may differ
between firmware versions, but the discovery methodology (systematic 0xE4
probing of the full 64KB space) remains applicable. The firmware version
string at XRAM \texttt{0x07F0} can help identify the specific variant.

Station-Drivers~\cite{station-drivers-asm} maintains a firmware archive
for ASM2362 variants, which could be used to identify XRAM layout
differences across firmware versions.

\subsubsection{Other Bridge Chipsets}

Other bridge chipsets (JMicron JMS583, Realtek RTL9210) use different
architectures and may or may not expose equivalent internal memory access.
The JMS583 has its own vendor commands but with different semantics. An
equivalent technique for these bridges would require separate reverse
engineering efforts.

\subsubsection{Other NVMe Controllers}

The Sanitize Block Erase approach is standard NVMe and should work on any
controller that supports it (check SANICAP). However, whether Sanitize
clears controller-specific FTL protection modes may vary by firmware
implementation. Controllers from different vendors (Samsung, Western
Digital, Micron) may respond differently.

\subsection{Ethical Considerations}

The XRAM injection technique bypasses security controls in the USB bridge
firmware. While our use case---recovering a personally owned drive---is
benign, the same technique could potentially be misused:

\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{Forensic sanitization bypass}: Sanitize commands sent
    through USB bridges may be silently dropped, meaning drives sanitized
    via USB are not reliably erased. Our technique demonstrates this gap
    but also provides a way to actually deliver the sanitize command.
  \item \textbf{Unauthorized command injection}: In shared USB
    infrastructure (e.g., USB hubs in labs or data centers), an attacker
    with access to the USB bus could inject NVMe commands into connected
    drives.
  \item \textbf{Firmware exploitation}: The absence of firmware signature
    verification~\cite{cyrozap-usb-pcie} on the ASM2362 means that
    arbitrary firmware could be loaded via 0xE3, potentially creating
    persistent USB bridge malware.
\end{itemize}

We note that the ASM2362's security posture was already weak (no signature
verification, XRAM exposed, UART debug available), making these concerns
largely pre-existing rather than novel. However, documenting these
capabilities carries responsibility. We have focused on the recovery use
case and provided safety mechanisms (dry-run default, address validation)
to minimize accidental misuse.

\subsection{Lessons Learned}

Several lessons from this investigation generalize beyond the specific
recovery scenario:

\begin{enumerate}[leftmargin=*]
  \item \textbf{USB bridges are not transparent tunnels.} They have their
    own firmware with opinions about which commands should pass through.
    Always verify actual command delivery at the target device, not just
    SCSI-layer status.
  \item \textbf{Silent drops are worse than errors.} The bridge returning
    SCSI GOOD for blocked commands misdirected weeks of debugging effort.
    An explicit error would have immediately pointed to the whitelist.
  \item \textbf{Vendor SCSI commands are the real power.} The 0xE6
    passthrough is the ``official'' API, but 0xE4/0xE5/0xE8 provide raw
    hardware access that the firmware's whitelist cannot filter.
  \item \textbf{Read the firmware RE before writing code.} cyrozap's
    documentation would have saved the entire 0xE6 Format/Sanitize
    implementation effort. In recovery/RE projects, start with what is
    known about the firmware.
  \item \textbf{Queue parameters must be discovered empirically.} Assumed
    queue depths and base addresses from documentation may not match the
    actual firmware configuration.
\end{enumerate}

\subsection{Limitations}

\begin{enumerate}[leftmargin=*]
  \item \textbf{BOT mode required}: The vendor SCSI commands fail under UAS,
    requiring a driver switch that may not be possible in all environments
    (e.g., macOS, some embedded Linux systems).
  \item \textbf{Bridge-specific XRAM layout}: Queue base addresses, depths,
    and MMIO register locations are specific to the ASM2362 firmware version
    and must be discovered per-device via systematic probing.
  \item \textbf{Single-byte writes}: The 0xE5 command writes one byte per
    SCSI transaction, making 64-byte SQE injection slow (approximately
    2--3 seconds for write plus verification).
  \item \textbf{No PCIe memory reads}: The TLP engine does not support
    non-posted reads, preventing direct access to NVMe controller registers
    beyond what is mapped into XRAM.
  \item \textbf{Data destruction}: Sanitize Block Erase destroys all data
    on the drive. This technique recovers the \emph{drive hardware}, not
    the \emph{data}. If data recovery is the goal, professional services
    with PC-3000 SSD are still required.
  \item \textbf{Controller-specific behavior}: The finding that only
    Sanitize (not Format) clears FTL protection is specific to the Phison
    PS5012-E12. Other controllers may behave differently.
  \item \textbf{Root access required}: SG\_IO ioctls require
    \texttt{CAP\_SYS\_RAWIO} or root privileges.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{Conclusion}
\label{sec:conclusion}

We have presented a novel technique for recovering write-protected NVMe SSDs
through USB bridge XRAM command injection. By exploiting the ASMedia ASM2362's
vendor-specific SCSI commands for direct XRAM access, we bypassed the bridge
firmware's NVMe opcode whitelist and successfully delivered a Sanitize Block
Erase command to a Phison PS5012-E12 controller that had been in read-only
mode for over a month following FTL corruption from a Windows hibernate
power loss event.

The key technical insights are:

\begin{enumerate}[leftmargin=*,nosep]
  \item The NVMe Admin Submission Queue resides in bridge XRAM and can be
    written via vendor SCSI commands (0xE5), bypassing the opcode whitelist.
  \item The PCIe TLP generation engine can ring the SQ Tail Doorbell
    without disconnecting USB, using a posted memory write TLP.
  \item Vendor SCSI commands (0xE4/0xE5/0xE8) require Bulk-Only Transport
    (BOT) mode, not UAS; this is a prerequisite that is not documented in
    the bridge specifications.
  \item The Admin SQ depth and CQ base address must be empirically
    determined, as they differ from the naive interpretation of XRAM
    region sizes (depth 4, not 8; CQ at \texttt{0xBC00}, not
    \texttt{0xB800}).
  \item Sanitize Block Erase, not Format NVM, is required to clear FTL
    write protection on the Phison PS5012-E12, despite Format being
    accepted and reported as successful.
  \item SMART Critical Warning bit 3 is not the sole indicator of
    firmware read-only mode; the controller can enter protection mode
    without setting this flag.
\end{enumerate}

The open-source implementation (\texttt{asm2362-tool}, approximately 4,000
lines of Zig with 35 unit tests) and this documentation provide a reusable
framework for NVMe SSD recovery through ASMedia USB bridges. The technique
potentially saves drives that would otherwise require expensive professional
recovery services or physical disassembly for native M.2 connection.

Future work includes extending the XRAM mapping to other ASMedia firmware
versions, investigating equivalent techniques for JMicron and Realtek
bridges, and developing automated queue parameter discovery that does not
require manual XRAM probing.

% ── Acknowledgments ──────────────────────────────────────────────────
\section*{Acknowledgments}

The author thanks cyrozap for the foundational ASM236x reverse engineering
work~\cite{cyrozap-usb-pcie} without which this technique would not have
been possible, the smartmontools developers~\cite{smartmontools} for the
reference \texttt{sntasmedia} implementation that confirmed the whitelist
limitation, and the usbdev.ru community for documenting Phison controller
recovery procedures and maintaining firmware archives.

% ── References ────────────────────────────────────────────────────────
\bibliographystyle{plain}
\bibliography{references}

% ══════════════════════════════════════════════════════════════════════
\appendix

\section{Hardware Stack Diagram}
\label{app:stack}

\begin{lstlisting}[style=hex,caption={Hardware stack (host to media)}]
+-------------------------------------------+
| Linux Host (Rocky Linux 10)               |
|   SG_IO ioctl -> /dev/sda                 |
|   Driver: usb-storage (BOT mode)          |
+--------------------+----------------------+
                     | USB 3.1 Gen 2 (10Gbps)
+--------------------+----------------------+
| ASMedia ASM2362 USB-NVMe Bridge           |
|   VID:PID = 174c:2362                     |
|   CPU: 8051 @ 114.3 MHz                   |
|   XRAM: 64KB (Admin SQ at 0xB000)        |
|   Firmware whitelist: 0x02, 0x06 only     |
|                                           |
| BYPASS PATH (our technique):              |
|   0xE4 -> Read XRAM (any address)         |
|   0xE5 -> Write XRAM (1 byte/cmd)         |
|   0xB210-B296 -> PCIe TLP engine          |
|     -> Memory write to BAR0+0x1000        |
+--------------------+----------------------+
                     | PCIe Gen 3 x2
+--------------------+----------------------+
| NVMe SSD Controller: Phison PS5012-E12    |
|   VID: 0x1E4B  FW: H211011a              |
|   OACS: 0x0006 (Format + FW Download)     |
|   SANICAP: 0x00000002 (Block Erase only)  |
|   Admin SQ: 4 entries (depth=4)           |
|   BAR0: 0x00D00000                        |
|   State: FTL write-protected (pre-fix)    |
+--------------------+----------------------+
                     | 8-channel, 3D TLC NAND
+--------------------+----------------------+
| Silicon Power SPCC M.2 PCIe SSD           |
|   Capacity: 256 GB (500M sectors)         |
|   Available Spare: 1%                     |
|   Media Errors: 292                       |
|   State: Read-only (FTL corrupted)        |
+-------------------------------------------+
\end{lstlisting}

\section{Complete Injection Workflow}
\label{app:workflow}

\begin{lstlisting}[style=hex,caption={Injection workflow flowchart},label=fig:workflow]
[1. Switch to BOT mode]
    |  modprobe usb-storage quirks=174c:2362:u
    v
[2. Probe bridge (0xE4 read at 0x0000)]
    |  Verify ASM2362 detected
    |  Read firmware version at 0x07F0
    v
[3. Read Admin SQ (0xE4, 256B at 0xB000)]
    |  Parse 4 x 64-byte SQ entries
    |  Identify target slot (0-3)
    |  Note: all 4 slots occupied at boot
    v
[4. Craft NVMe SQE]
    |  Sanitize: OPC=0x84, SANACT=2
    |  Set CID=0x4242 for CQ correlation
    |  NSID=0 (controller-wide)
    |  PRP1=PRP2=0 (no data transfer)
    v
[5. Write SQE to XRAM]
    |  64 x 0xE5 SCSI commands
    |  Each: write byte + readback verify
    |  Abort on any verification failure
    |  ~2-3 seconds total
    v
[6. Full readback verification]
    |  0xE4 read 64 bytes at slot address
    |  Compare against expected SQE bytes
    |  std.mem.eql() exact match required
    v
[7. DRY-RUN DECISION POINT]
    |
    +--[--dry-run]-> STOP (safe inspection)
    |
    +--[--force]--> CONTINUE
    v
[8. Read BAR0 via PCIe config read TLP]
    |  Type 1 cfg read: bus=1 dev=0 fn=0
    |  Register offset 0x10
    |  Result: BAR0 = 0x00D00000
    v
[9. Ring doorbell: PCIe memory write TLP]
    |  Write tail=1 to 0x00D01000
    |  TLP header at XRAM 0xB210
    |  TLP data at XRAM 0xB220
    |  Trigger at XRAM 0xB254
    |  Posted write: no completion
    |  USB connection stays alive
    v
[10. Read Admin CQ (0xE4 at 0xBC00)]
    |  Find entry with CID=0x4242
    |  Check: SCT=0, SC=0 -> SUCCESS
    |  Phase tag=1 -> valid new entry
    v
[11. CPU reset (0xE8 type=0)]
    |  8051 restarts
    |  NVMe controller re-initializes
    |  Drive re-enumerates on USB bus
    v
[12. Post-recovery verification]
    |  dd write 1MB random data
    |  dd read back + md5sum compare
    |  SMART: SSTAT=0x0001 (complete)
    |  identify: controller responds
    v
[SUCCESS: Drive recovered]
\end{lstlisting}

\section{NVMe Admin Opcodes Reference}
\label{app:opcodes}

\begin{table}[htbp]
\centering
\caption{NVMe Admin opcodes relevant to recovery}
\label{tab:opcodes-ref}
\footnotesize
\begin{tabular}{@{}clc@{}}
\toprule
\textbf{Opcode} & \textbf{Command} & \textbf{CDW10 Key Bits} \\
\midrule
\texttt{0x00} & Delete I/O SQ & QID \\
\texttt{0x01} & Create I/O SQ & QID, QSIZE \\
\texttt{0x02} & Get Log Page  & LID, NUMD \\
\texttt{0x04} & Delete I/O CQ & QID \\
\texttt{0x05} & Create I/O CQ & QID, QSIZE \\
\texttt{0x06} & Identify      & CNS \\
\texttt{0x09} & Set Features  & FID, SV \\
\texttt{0x0A} & Get Features  & FID, SEL \\
\texttt{0x80} & Format NVM    & LBAF, SES \\
\texttt{0x81} & Security Send & SECP, SPSP \\
\texttt{0x82} & Security Recv & SECP, SPSP \\
\texttt{0x84} & Sanitize      & SANACT \\
\bottomrule
\end{tabular}
\end{table}

\section{XRAM Hex Dump: Admin SQ Pre-Recovery}
\label{app:hexdump}

The Admin SQ state before injection, showing 4 occupied entries from the
bridge's initialization sequence:

\begin{lstlisting}[style=hex,caption={Admin SQ at 0xB000 (pre-injection, 256 bytes)}]
     ; Slot 0: Set Features (FID=0x07, Num Queues)
B000: 09 00 01 00 00 00 00 00  ; CDW0: OPC=0x09 CID=1
B008: 00 00 00 00 00 00 00 00  ; Reserved
B010: 00 00 00 00 00 00 00 00  ; MPTR
B018: 00 00 00 00 00 00 00 00  ; PRP1
B020: 00 00 00 00 00 00 00 00  ; PRP2
B028: 07 00 00 00 FE 00 FE 00  ; CDW10=0x07 CDW11
B030: 00 00 00 00 00 00 00 00  ; CDW12-13
B038: 00 00 00 00 00 00 00 00  ; CDW14-15

     ; Slot 1: Set Features (second)
B040: 09 00 02 00 00 00 00 00  ; CDW0: OPC=0x09 CID=2
B048: 00 00 00 00 00 00 00 00
...  (similar structure)

     ; Slot 2: Create I/O CQ
B080: 05 00 03 00 00 00 00 00  ; CDW0: OPC=0x05 CID=3
B088: 00 00 00 00 00 00 00 00
...  (queue parameters in CDW10-11)

     ; Slot 3: Create I/O SQ
B0C0: 01 00 04 00 00 00 00 00  ; CDW0: OPC=0x01 CID=4
B0C8: 00 00 00 00 00 00 00 00
...  (queue parameters in CDW10-11)

     ; Slot 4: Beyond queue boundary (empty)
B100: 00 00 00 00 00 00 00 00
B108: 00 00 00 00 00 00 00 00
...  (all zeros through 0xB1FF)
\end{lstlisting}

The four entries represent the bridge firmware's NVMe initialization
sequence: configure queue counts (Set Features FID=0x07), configure
additional features, create I/O completion queue (CQ ID=1), and create
I/O submission queue (SQ ID=1). All four valid slots are occupied, which
is why the injection must overwrite an existing entry at \texttt{slot=0}
and set the doorbell tail to 1 to trigger processing of the new command.

\section{PCIe TLP Register Map}
\label{app:tlp-regs}

\begin{table}[htbp]
\centering
\caption{ASM2362 PCIe TLP engine XRAM registers}
\label{tab:tlp-regs}
\footnotesize
\begin{tabular}{@{}llp{3.2cm}@{}}
\toprule
\textbf{Address} & \textbf{Size} & \textbf{Function} \\
\midrule
\texttt{0xB210} & 4B & TLP header DW0 (fmt\_type, length) \\
\texttt{0xB214} & 4B & TLP header DW1 (byte enables) \\
\texttt{0xB218} & 4B & TLP header DW2 (address) \\
\texttt{0xB220} & 4B & TLP write data (u32 BE) \\
\texttt{0xB224} & 12B & TLP completion header \\
\texttt{0xB254} & 1B & Operation trigger (write 0x0F) \\
\texttt{0xB284} & 1B & Config request status \\
\texttt{0xB296} & 1B & Control/Status Register \\
\bottomrule
\end{tabular}
\end{table}

\noindent CSR (\texttt{0xB296}) bit definitions:
\begin{description}[nosep,style=unboxed,leftmargin=0.5cm]
  \item[Bit 0] Timeout flag (write 0x01 to clear)
  \item[Bit 1] Completion done (for non-posted TLPs)
  \item[Bit 2] Ready to send (poll before sending)
\end{description}

\noindent TLP type codes used:
\begin{description}[nosep,style=unboxed,leftmargin=0.5cm]
  \item[\texttt{0x00}] Memory Read (3DW, no data) -- \emph{does not work}
  \item[\texttt{0x04}] Config Read Type 0
  \item[\texttt{0x05}] Config Read Type 1 (used for BAR0)
  \item[\texttt{0x40}] Memory Write (3DW, with data) -- \emph{used for doorbell}
  \item[\texttt{0x44}] Config Write Type 0
  \item[\texttt{0x45}] Config Write Type 1
\end{description}

\end{document}
